\section{Methods}
\label{sec:methods}

\subsection{Notation}
\label{sec:notation}

Throughout, $x\in\mathbb{R}^{L}$ is the stored single-period wavetable (raw engine input).
$y=\Theta(x;\theta)$ is the engine-baked cycle after seam ops and optional residual cells with parameters $\theta$.
$r^{\star}\in\mathbb{R}^{L}$ is a procedural ideal sibling from the same seed and family with the open-wrap cliff withheld.
Studio-clean recordings are not used as $r^{\star}$.
Tiling length $N$ (typically $N{=}16$) forms $y_{\mathrm{tiled}}=\mathrm{Tile}(y,N)$ and $r^{\star}_{\mathrm{tiled}}=\mathrm{Tile}(r^{\star},N)$.
Seam width $W$ (code: \texttt{SEAM\_W}, typically $8$) is the neighborhood edited by classical fade and polish ops.
$R\in[0,1]$ is the prolonged residual score defined below ($1$~=~best).
These symbols are fixed for the residual objective. They follow a different convention from Noise2Noise clean/noisy pairs.

\begin{table}[t]
  \centering
  \caption{Core symbols used in Methods.}
  \label{tab:symbols}
  \setlength{\tabcolsep}{3pt}
  \begin{tabular}{@{}cl@{}}
    \toprule
    Symbol & Meaning \\
    \midrule
    $L$ & Cycle length (samples per period) \\
    $x$ & Raw engine wavetable period \\
    $y=\Theta(x;\theta)$ & Baked cycle after seam ops / residual cell \\
    $r^{\star}$ & Ideal sibling (open-wrap cliff withheld) \\
    $N$ & Prolong tiling count \\
    $W$ & Seam neighborhood width \\
    $R$ & Prolonged residual score in $[0,1]$ (1~=~best) \\
    $\theta$ & Continuous bake / residual parameters \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Ideal tile construction}
\label{sec:ideal-tile}

The ideal sibling is generated by the same procedural draw as the engine input (frequency, phase, mid-cycle noise policy) with the open-wrap cliff amplitude set to zero at both ends.
Engine and ideal therefore share the mid-cycle content of one seed draw.
They differ by the seam cliff (and any seam-boosted noise policy applied only to the engine path).
This sibling relationship is what makes unsupervised ranking possible without paired studio recordings~\cite{lehtinen2018,kashyap2021n2n}.

\subsection{Residual score $R$}
\label{sec:residual-score}

\textbf{Every} method---no-bake, DualCosine, FIR, poly, VA residuals, neural favorite---is scored the same way:
apply an operator $\Theta$ to the cracked engine period $x$ to obtain $y=\Theta(x)$, then compare the tiled $y$ to the \emph{same} ideal sibling $r^{\star}$ under~\eqref{eq:R}.
The ideal sibling is the scoring target (open-wrap cliff withheld).
It is \emph{not} a method output and it is \emph{not} the no-bake baseline.

With tiling length $N$,
\begin{equation}
\label{eq:R}
R=\mathrm{clamp}\!\left(1-\frac{\mathrm{rms}(y_{\mathrm{tiled}}-r^{\star}_{\mathrm{tiled}})}{\max(\mathrm{rms}(r^{\star}_{\mathrm{tiled}}),\varepsilon)},\,0,\,1\right),
\end{equation}
where $\mathrm{rms}(z)=\sqrt{\frac{1}{|z|}\sum_i z_i^2}$ and $\varepsilon>0$ avoids division by zero.
By construction $R\in[0,1]$, $R{=}1$ when $y_{\mathrm{tiled}}=r^{\star}_{\mathrm{tiled}}$, and (when pre-clamp scores lie in $(0,1)$) $R$ is strictly decreasing in residual RMS for fixed ideal RMS.
A method ``beats'' another iff its $R$ is higher on the same $(x,r^{\star})$ draw---equivalently, iff its bake is closer to the ideal sibling.
Optional soft shape gates (when enabled) multiply $R$ by factors that down-weight mid-cycle damage relative to seam-local error.
We do not claim convergence guarantees for the hybrid GA{+}PPO{+}PBT outer loop, nor a general wrap-closure theorem for arbitrary bake cells.
Empirical gains are reported under the frozen protocol.
Outer-loop cost scales as population size $\times$ proposals per iteration $\times$ fit steps $\times$ batch forward cost (order $10^{3}$--$10^{4}$ arch evaluations at the reported $5$k gate). We report measured budgets, not a convexity claim.

\begin{remark}[Noise2Noise motivation]
\label{rem:n2n}
Ranking without studio-clean pairs is possible when the ideal is a procedural sibling of the engine draw~\cite{lehtinen2018,kashyap2021n2n}.
The remark is a design philosophy for procedural siblings.
\end{remark}

\begin{algorithm}[t]
\caption{ResidualScore}
\label{alg:residual}
\begin{algorithmic}[1]
\Require Ideal cycle $r^{\star}$, baked cycle $y$, periods $N$
\State $I \gets \mathrm{Tile}(r^{\star}, N)$
\State $Y \gets \mathrm{Tile}(y, N)$
\State $r_{\mathrm{rms}} \gets \mathrm{RMS}(Y-I)$
\State $i_{\mathrm{rms}} \gets \max(\mathrm{RMS}(I), \varepsilon)$
\State \Return $\mathrm{clamp}(1 - r_{\mathrm{rms}}/i_{\mathrm{rms}},\,0,\,1)$
\end{algorithmic}
\end{algorithm}

\subsection{Bake operator and baselines}
\label{sec:bake}

A \textbf{bake cell} $\Theta(\cdot;\theta)$ composes classical seam ops (DualCosine, polish, soft seam, FIR blends) with optional tiny residual cells (MLP/FIR/U-Net-lite/attention-lite) and optional MoE soft gates over experts.
Bake parameters $\theta$ live in a bounded box. Architecture strings are discrete graphs over those ops.
Separate classical controls apply cycle-local BLIT/BLEP, PolyBLEP, and BLAMP residual corrections at the wrap on the cracked engine tile (same interface as DualCosine / FIR; not a full oscillator rewrite).
Plain-language roles: BLIT/BLEP and PolyBLEP correct a \emph{value} step at wrap.
BLAMP corrects a \emph{slope}/corner jump.
DualCosine fades endpoints without a BLEP residual.
They are scored and illustrated in Section~\ref{sec:va-seam-baselines}.

\paragraph{No-bake (passthrough) control.}
\label{sec:no-bake}
The \textbf{no-bake} baseline is the unrepaired cracked engine: $\Theta_{\mathrm{no\text{-}bake}}(x)=x$, then $R=\mathrm{ResidualScore}(r^{\star},x,N)$.
It is the engine period \emph{as stored}, scored against the ideal sibling.
It is \emph{not} the ideal sibling itself, not a residual-network skip, not a learned identity map, and not a repair method.
On mild cliffs residual mass $\ll$ ideal RMS under~\eqref{eq:R}, so no-bake $R$ can sit near $0.97$ while the wrap click remains audible---a reference floor/ceiling for ``do nothing,'' not a perceptual wrap score.
Released JSON artifacts may still use the legacy key \texttt{identity} for this control; manuscript tables and prose use \emph{no-bake}.

\paragraph{Classical (non-AI) comparison board.}
\label{sec:classical-board}
Learned / searched methods are always ranked against the full classical board on absolute $R$ (same $r^{\star}$):
no-bake, DualCosine / cosine fade, \texttt{seam\_fir3}, classic quadratic, hann/crossfade/soft fades, detrend/ensemble, poly seam, and cycle-local BLIT/BLEP / PolyBLEP / BLAMP where scored.
The optimization objective is to \emph{maximize} prolonged $R$ toward the ideal sibling (best attainable $R{=}1$ under~\eqref{eq:R}); higher $R$ is strictly better.
DualCosine appears in PPO as a \emph{constant centering} term ($\rho{=}R-R_{\mathrm{DualCosine}}$) so advantages are roughly zero-mean early in search---it is not the reward target and not the sole classical comparator.
When tables show $\Delta R$ vs DualCosine, that is a reporting gap vs one classical bake, not the definition of reward.

\begin{algorithm}[t]
\caption{DualCosine baseline}
\label{alg:dualcosine}
\begin{algorithmic}[1]
\Require Engine cycle $x$
\State Fade both ends of $x$ with raised-cosine windows of width $W$
\State \Return $\mathrm{ResidualScore}(r^{\star},\,x_{\mathrm{faded}},\,N)$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
\caption{No-bake (passthrough) control}
\label{alg:no-bake}
\begin{algorithmic}[1]
\Require Engine cycle $x$ (open-wrap cliff present)
\State \Return $\mathrm{ResidualScore}(r^{\star},\,x,\,N)$ \Comment{no operator: $x\mapsto x$}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
\caption{MoE soft gating over bake experts}
\label{alg:moe}
\begin{algorithmic}[1]
\Require Cycle $x$, expert cells $\{E_k\}_{k=1}^{K}$, gate logits $g\in\mathbb{R}^{K}$
\State $w \gets \mathrm{softmax}(g)$
\State $y \gets \sum_{k=1}^{K} w_k\, E_k(x)$
\State \Return $y$
\end{algorithmic}
\end{algorithm}

\subsection{Search space}
\label{sec:search-space}

Each genome encodes: ordered op list (fade, polish, FIR, pin, residual write), depth (number of composed blocks), residual cell type (none, MLP, FIR, U-Net-lite, attn-lite, LSTM-lite), optional MoE expert count $K\in\{2,3,4\}$, and continuous $\theta$ in box bounds used by the CUDA search runner (fade widths, FIR taps, residual scales).
The searchable bake vocabulary includes a width-capped 1--2 layer LSTM block over the seam window or full cycle (distinct from the fixed supervised seq-LSTM SOTA baseline).
The live hybrid tag is \texttt{PPO+GA+PBT+NAS+depth+MoE} with LSTM enabled in the block graph.

\subsection{Training / fit inner loop}
\label{sec:fit}

\begin{algorithm}[t]
\caption{FitCell (inner loop)}
\label{alg:fit}
\begin{algorithmic}[1]
\Require Architecture cfg, fit steps $T$, batch size $B$
\State cell $\gets \mathrm{SeamCell}(\mathrm{cfg})$; opt $\gets \mathrm{Adam}(\mathrm{cell.params})$
\State prev $\gets \mathrm{null}$; patience $\gets 0$
\For{$t = 1,\ldots,T$}
  \State ideal, eng $\gets \mathrm{MakeBatch}(B)$
  \State out $\gets \mathrm{ApplyOps}(\mathrm{eng},\,\mathrm{cell},\,\mathrm{cfg.ops})$
  \State $R \gets \mathrm{mean}\,\mathrm{ResidualScore}(\mathrm{ideal},\,\mathrm{out})$
  \State loss $\gets 1-R$
  \If{finite(loss)} AdamStep(opt, loss)
  \EndIf
  \If{prev $\neq$ null and $|\mathrm{prev}-R|/\max(|\mathrm{prev}|,\varepsilon) < 10^{-4}$}
    \State patience $\gets$ patience $+1$; \textbf{if} patience $\ge 3$ \textbf{then break}
  \Else
    \State patience $\gets 0$
  \EndIf
  \State prev $\gets R$
\EndFor
\State \Return cell, $R$
\end{algorithmic}
\end{algorithm}

\subsection{Hybrid outer loop}
\label{sec:hybrid}

Each iteration rotates among branches: PPO, GA, PBT, NAS, and combo.
Selection and FitCell are always by absolute residual $R$ (maximize toward the ideal sibling; loss $1-R$).
The reward is larger when $R$ is closer to the best possible value ($R{\to}1$).

\paragraph{Near-ceiling $R$ and why reward / search HPs matter.}
\label{sec:reward-shaping}
On the sine{+}cliff generator, no-bake already sits near $R\approx 0.97$, and strong classical / searched bakes live in a narrow band toward $1$ (often $0.99$--$0.991$).
Absolute gaps of order $10^{-3}$--$10^{-2}$ are scientifically real (closer to $r^{\star}$) but tiny as raw PPO/GA signals: without shaping, advantages collapse and hyperparameters (clip, entropy, mutation rate, population size, fit steps, learning rate) dominate whether the outer loop can climb the last percent.
We therefore treat \textbf{reward shaping and search hyperparameters as first-class tunables} alongside bake architecture---not as fixed afterthoughts.
In the present hybrid, PPO uses a searchable shaping mode among $\{\texttt{abs\_r},\,\texttt{vs\_dualcosine},\,\texttt{vs\_nobake},\,\texttt{neglog\_gap}\}$ (PBT-mutated; default $\rho = R - R_{\mathrm{DualCosine}}$).
DualCosine centering is one shaping choice, not the objective.
PBT mutates fit / PPO hyperparameters and reward mode in-population; plateau adaptation retunes branch mix and depth/MoE bias when champions stall.
The matched $5$k meta-approach comparison (Section~\ref{sec:meta-compare}) is the sole GPU experimentation vehicle for outer-loop family, HP, and reward-shaping sweeps under a shared FitCell budget (no concurrent long hybrid campaign).

\begin{algorithm}[t]
\caption{GA tournament step}
\label{alg:ga}
\begin{algorithmic}[1]
\Require Population $P$, tournament size $k$, crossover rate $p_c$, mutation rate $p_m$
\State Sample $k$ genomes; parents $\gets$ top-2 by $R$
\State child $\gets$ Crossover(parents) with prob.\ $p_c$, else clone elite
\State Mutate(child) with rate $p_m$ (ops, depth, cell type, $\theta$)
\State Fit and evaluate child; insert into $P$ (replace worst if full)
\State \Return updated $P$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
\caption{PPO architecture update}
\label{alg:ppo}
\begin{algorithmic}[1]
\Require Actor--critic $\pi_\phi$, clip $\epsilon$, experience buffer of mutate actions
\State Sample action $a$ editing ops / hyperparameters from $\pi_\phi$
\State Fit proposed cfg; observe centered advantage $\rho = R - R_{\mathrm{DualCosine}}$ \Comment{objective remains maximize $R$}
\State Advantage $\hat{A}$ from critic; ratio $r_\phi = \pi_\phi(a)/\pi_{\phi_{\mathrm{old}}}(a)$
\State Update $\phi$ with clipped surrogate $\min(r_\phi\hat{A},\,\mathrm{clip}(r_\phi,1-\epsilon,1+\epsilon)\hat{A})$
\State \Return updated $\pi_\phi$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
\caption{PBT exploit--mutate}
\label{alg:pbt}
\begin{algorithmic}[1]
\Require Population $P$, exploit quantile $q$, mutate scale $\sigma$
\State Rank $P$ by $R$; for each bottom member, copy hyperparameters from a top-$q$ elite
\State Perturb copied hyperparameters by multiplicative noise of scale $\sigma$
\State Refit briefly; write back into $P$
\State \Return updated $P$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
\caption{HybridMetaSearch}
\label{alg:hybrid}
\begin{algorithmic}[1]
\Require Iteration budget $T_{\mathrm{out}}$, population size $n$
\State $P \gets \mathrm{InitPopulation}(n)$; $\pi \gets \mathrm{ActorCritic}()$; champ $\gets$ null; boredom $\gets 0$
\For{$it = 1,\ldots,T_{\mathrm{out}}$}
  \State branch $\gets \mathrm{Rotate}(\mathrm{ppo},\mathrm{ga},\mathrm{pbt},\mathrm{nas},\mathrm{combo})$
  \State cfg, hp $\gets \mathrm{Propose}(\mathrm{branch},\,P,\,\pi)$
  \State cell, $R_{\mathrm{fit}} \gets \mathrm{FitCell}(\mathrm{cfg},\,\mathrm{hp.fit\_steps})$
  \State $R_{\mathrm{eval}} \gets \mathrm{EvalCell}(\mathrm{cell})$
  \State $R \gets \tfrac12 R_{\mathrm{fit}} + \tfrac12 R_{\mathrm{eval}}$ (+ optional depth/MoE bonus)
  \State UpdatePopulation($P$, cfg, $R$); PPOUpdate($\pi$, centered $\rho = R - R_{\mathrm{DualCosine}}$)
  \If{$R > \mathrm{champ}.R$} champ $\gets$ (cfg, cell, $R$); boredom $\gets 0$
  \Else\ boredom $\gets$ boredom $+1$
  \EndIf
  \If{boredom $\ge 1000$} PlateauAdapt; boredom $\gets 0$
  \EndIf
\EndFor
\State \Return champ
\end{algorithmic}
\end{algorithm}

\subsection{Meta-learning approach comparison}
\label{sec:meta-compare}

Beyond the hybrid outer loop, we compare five additional outer-loop families under a matched $5{,}000$-evaluation budget, shared search seed \texttt{1902771841}, and the same FitCell / prolonged-$R$ protocol (LSTM and xLSTM included in the searchable bake vocabulary).
Table and figure legends use the short names below (artifact folders keep stable code keys; see \texttt{NOMENCLATURE.md}):
\begin{itemize}
  \item \textbf{Random NAS} (\texttt{random}): independent uniform draws over bake graphs and fit hyperparameters (control).
  \item \textbf{Cont.\ CMA-ES} (\texttt{cmaes}): diagonal covariance-matrix adaptation over a continuous encoding of depth, width, cell kind, block logits, and fit hyperparameters~\cite{hansen2016cmaes}.
  \item \textbf{Arch REINFORCE} (\texttt{reinforce}): vanilla policy gradient over discrete architecture-edit actions (no clipped surrogate; contrast to PPO~\cite{schulman2017,zoph2017nas}).
  \item \textbf{Aging evolution} (\texttt{aging\_evo}): regularized evolution with oldest-member replacement after tournament parent selection~\cite{real2017large,real2019}, distinct from the hybrid's non-aging tournament GA branch.
  \item \textbf{TPE Bayes NAS} (\texttt{tpe}): lightweight tree-structured Parzen estimator over discrete bake categoricals in the spirit of practical Bayesian optimization~\cite{snoek2012}.
\end{itemize}
We also run \textbf{Ours (hybrid GA--PPO)} (\texttt{hybrid\_lstm}): the multi-branch hybrid loop for $5{,}000$ evaluations with LSTM/xLSTM in the block vocabulary and online reward-mode / fit--PPO hyperparameter co-tuning, as the proposed matched-budget arm.
Results appear in Table~\ref{tab:meta-approaches} and Figure~\ref{fig:meta-approach-compare}.

\begin{remark}[Trial complexity]
\label{rem:complexity}
Per outer iteration the dominant cost is $O(T_{\mathrm{fit}}\cdot C_{\mathrm{fwd}})$ for fit steps times one forward bake.
Bake-width caps (tiny residual cells) keep $C_{\mathrm{fwd}}$ far below full-band Demucs/DiffWave forwards screened in Related Work.
\end{remark}

\input{subsections/arch_diagram}

\subsection{Hyperparameters}
\label{sec:hyperparams}

Table~\ref{tab:hyperparams} lists the CUDA search defaults used for the $5$k-gate freeze reported in Results.
Under the near-ceiling residual regime (Section~\ref{sec:reward-shaping}), these outer-loop and fit hyperparameters are co-tuned with architecture; PBT and plateau adaptation continue to mutate several of them online.
Canonical companion pseudocode lives in \texttt{docs/PSEUDOCODE.md} (mirrored Algorithms~\ref{alg:residual}--\ref{alg:hybrid}).

\begin{table}[t]
  \centering
  \small
  \caption{Hybrid search hyperparameters (RTX~3090 DenoiseOpt CUDA search runner / \texttt{denoise\_meta\_evo.py} defaults). Co-tuned with reward shaping under near-ceiling $R$ (Section~\ref{sec:reward-shaping}).}
  \label{tab:hyperparams}
  \setlength{\tabcolsep}{3pt}
  \begin{tabular}{@{}lp{0.48\columnwidth}@{}}
    \toprule
    Parameter & Value \\
    \midrule
    Population size $n$ & 12 \\
    GA tournament $k$ & 3 \\
    GA crossover fraction $p_c$ & $0.5$ (else clone elite parent) \\
    GA mutation & $\ge 1$ mutate after inherit ($+50\%$ second mutate) \\
    PPO clip $\epsilon$ & $0.2$ (default; searchable $\{0.1,0.2,0.25,0.3\}$) \\
    Adam learning rate (fit) & $3{\times}10^{-3}$ (default; PBT-searchable) \\
    Adam learning rate (policy) & $3{\times}10^{-4}$ \\
    Entropy coefficient (PPO) & $0.02$ (default; searchable $\approx 0.005$--$0.06$) \\
    PBT exploit / mutate & elite fraction $0.25$; multi-step arch+hp mutate \\
    MoE routing & \texttt{sequential} or \texttt{moe\_parallel} when enabled \\
    Plateau boredom threshold & $1000$ iters without champ improve \\
    Cycle length $L$ / tiles $N$ / seam $W$ & $256$ / $16$ / $8$ \\
    Fit steps / batch (default) & $24$ / $48$ \\
    Algo tag & \texttt{PPO+GA+PBT+NAS+depth+MoE} \\
    \bottomrule
  \end{tabular}
\end{table}
