%% DenoiseOpt paper v4 — arXiv-style double-column preprint
%% Template: article[twocolumn] + local arxiv-twocolumn.sty
%% See README.md for rationale and sources.
\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{arxiv-twocolumn}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{placeins}
\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}

\title{Hybrid Reinforcement Learning and Genetic Algorithms\\
for Meta-Learning Audio Denoising}
\author{%
  Julian~M.~Kleber\\[2pt]
  \normalsize
  ORCID:
  \href{https://orcid.org/0000-0001-5518-0932}{0000-0001-5518-0932}\\[1pt]
  Email:
  \href{mailto:julian.m.kleber@gmail.com}{\texttt{julian.m.kleber@gmail.com}}
}
\date{18 July 2026\\[2pt]
  \small Paper version \texttt{v4}
  (introduction + literature landscape; experiments ongoing)}
% Short running header (full title is too long for the column header)
\newcommand{\papershorttitle}{Hybrid RL+GA for Audio Denoising}

\begin{document}

% Full-width title / abstract / code banner, then two-column body
\twocolumn[{%
  \maketitle
  \begin{center}\footnotesize
    Code:
    \url{https://github.com/reeldemo/reelsynth},
    \url{https://github.com/reeldemo/denoise-opt-meta}
  \end{center}
  \begin{abstract}
    Periodic discontinuities in cyclic audio frames inject broadband crackle under wrap-around playback.
    We treat cycle-local seam restoration as an instance of audio signal denoising and search denoising operators with a hybrid meta-learning outer loop that interleaves genetic-algorithm population updates with reinforcement-learning architecture proposals, optionally with population-based exploit--mutate schedules and mixture-of-experts soft gates.
    Candidates are scored by a prolonged residual $R\in[0,1]$ (1~=~best) against an ideal multi-period tiling, without paired clean waveforms for every trial.
    This draft delivers the problem framing and a used-versus-screened literature landscape grounded on OA PDF analyses (Noise2Noise, PBT, Aging Evolution, ERL, ENAS/PPO, Wave-U-Net/Conv-TasNet, MoE, DDSP; Demucs/DiffWave/MAML/DARTS screened).
    A large GPU campaign (\texttt{PPO+GA+PBT+NAS}) is ongoing; final quantitative tables are deferred until the declared budget completes.
  \end{abstract}
  \keywords{audio denoising $\cdot$ neural architecture search $\cdot$ reinforcement learning $\cdot$ genetic algorithms $\cdot$ wavetable synthesis $\cdot$ meta-learning}
  \vspace{0.6em}
}]

\input{subsections/introduction}
\input{subsections/related_work}
\input{subsections/methods}
\input{subsections/experiments}
\input{subsections/results}
\input{subsections/discussion}
\input{subsections/limitations}
\input{subsections/conclusion}

\FloatBarrier

\begin{thebibliography}{99}

\bibitem{valimaki2006va} V.~V\"alim\"aki, A.~Huovilainen (2006).
\emph{Oscillator and Filter Algorithms for Virtual Analog Synthesis}.
Computer Music Journal, 30(2).

\bibitem{esqueda2016aliasing} F.~Esqueda, S.~Bilbao, V.~V\"alim\"aki (2016).
\emph{Aliasing Reduction in Clipped Signals}.
IEEE Transactions on Signal Processing.

\bibitem{esqueda2016blamp} F.~Esqueda, V.~V\"alim\"aki, S.~Bilbao (2016).
\emph{Rounding Corners with BLAMP}.
Proc.\ DAFx.

\bibitem{lehtinen2018} J.~Lehtinen et al.\ (2018).
\emph{Noise2Noise: Learning Image Restoration without Clean Data}.
ICML.\ arXiv:1803.04189.

\bibitem{engel2020} J.~Engel, L.~Hantrakul, C.~Gu, A.~Roberts (2020).
\emph{DDSP: Differentiable Digital Signal Processing}.
ICLR.\ arXiv:2001.04643.

\bibitem{purwins2019} H.~Purwins et al.\ (2019).
\emph{Deep Learning for Audio Signal Processing}.
IEEE J.\ Sel.\ Topics Signal Process.\ arXiv:1905.00078.

\bibitem{stoller2018} D.~Stoller, S.~Ewert, S.~Dixon (2018).
\emph{Wave-U-Net: A Multi-Scale Neural Network for End-to-End Audio Source Separation}.
ISMIR.\ arXiv:1806.03185.

\bibitem{jansson2017} A.~Jansson et al.\ (2017).
\emph{Singing Voice Separation with Deep U-Net Convolutional Networks}.
ISMIR.

\bibitem{luo2019} Y.~Luo, N.~Mesgarani (2019).
\emph{Conv-TasNet: Surpassing Ideal Time--Frequency Magnitude Masking for Speech Separation}.
IEEE/ACM TASLP.

\bibitem{elsken2019} T.~Elsken, J.~H.~Metzen, F.~Hutter (2019).
\emph{Neural Architecture Search: A Survey}.
JMLR.\ arXiv:1808.05377.

\bibitem{pham2018} H.~Pham et al.\ (2018).
\emph{Efficient Neural Architecture Search via Parameter Sharing}.
ICML.\ arXiv:1802.03268.

\bibitem{schulman2017} J.~Schulman et al.\ (2017).
\emph{Proximal Policy Optimization Algorithms}.
arXiv:1707.06347.

\bibitem{real2019} E.~Real, A.~Aggarwal, Y.~Huang, Q.~V.~Le (2019).
\emph{Regularized Evolution for Image Classifier Architecture Search}.
AAAI.\ arXiv:1802.01548.

\bibitem{jaderberg2017} M.~Jaderberg et al.\ (2017).
\emph{Population Based Training of Neural Networks}.
arXiv:1711.09846.

\bibitem{khadka2018} S.~Khadka, K.~Tumer (2018).
\emph{Evolution-Guided Policy Gradient in Reinforcement Learning}.
NeurIPS.\ arXiv:1805.07917.

\bibitem{shazeer2017} N.~Shazeer et al.\ (2017).
\emph{Outrageously Large Neural Networks: The Sparsely-Gated Mixture-of-Experts Layer}.
ICLR.\ arXiv:1701.06538.

\bibitem{snoek2012} J.~Snoek, H.~Larochelle, R.~P.~Adams (2012).
\emph{Practical Bayesian Optimization of Machine Learning Algorithms}.
NeurIPS.\ arXiv:1206.2944.

\bibitem{zhang2007} Q.~Zhang, H.~Li (2007).
\emph{MOEA/D: A Multiobjective Evolutionary Algorithm Based on Decomposition}.
IEEE Trans.\ Evolutionary Computation.

\bibitem{finn2017} C.~Finn, P.~Abbeel, S.~Levine (2017).
\emph{Model-Agnostic Meta-Learning for Fast Adaptation of Deep Networks}.
ICML.\ arXiv:1703.03400.\ \textit{(Screened.)}

\bibitem{liu2019} H.~Liu, K.~Simonyan, Y.~Yang (2019).
\emph{DARTS: Differentiable Architecture Search}.
ICLR.\ arXiv:1806.09055.\ \textit{(Screened.)}

\bibitem{dfossez2019} A.~D\'efossez et al.\ (2019).
\emph{Demucs: Deep Extractor for Music Sources}.
arXiv:1911.13254.\ \textit{(Screened.)}

\bibitem{dfossez2021} A.~D\'efossez (2021).
\emph{Hybrid Spectrogram and Waveform Source Separation}.
arXiv:2111.03600.\ \textit{(Screened.)}

\bibitem{kong2021} Z.~Kong et al.\ (2021).
\emph{DiffWave: A Versatile Diffusion Model for Audio Synthesis}.
ICLR.\ arXiv:2009.00713.\ \textit{(Screened.)}

\bibitem{pascual2017} S.~Pascual, A.~Bonafonte, J.~Serr\`a (2017).
\emph{SEGAN: Speech Enhancement Generative Adversarial Network}.
INTERSPEECH.\ \textit{(Screened.)}

\end{thebibliography}
\end{document}
