\section{Discussion}
\label{sec:discussion}

\paragraph{Where residual mass remains.}
Across the $100{,}000$-cycle procedural bench and lit-combo family stress tests, wrap error is not uniform.
The same generator families repeatedly dominate the residual budget: \texttt{nonlinear}, \texttt{combo}, \texttt{triple\_mix}, and \texttt{extreme\_overlay} show the weakest denoise / residual scores, while \texttt{harmonic\_fft} and \texttt{am\_fm} are comparatively easy.
\texttt{open\_wrap\_bias} can look strong on wrap-energy proxies yet still lag on prolonged residual. Cliff reduction and ideal-matched tiling are related objectives with different rankings.

\paragraph{No-bake $R$ and hard-cliff honesty.}
Canonical no-bake $R\approx 0.97$ looks ``easy'' only if one treats $R$ as a perceptual wrap score or confuses no-bake with the ideal sibling.
No-bake is the unrepaired engine scored against $r^{\star}$; the ideal withholds the cliff and is never a method output.
Under~\eqref{eq:R}, small residual RMS relative to ideal RMS yields high $R$ even when $|x_0-x_{L-1}|$ is large.
That same near-ceiling regime is why tiny $\Delta R$ still matters operationally: climbing $0.97\to 0.991$ is most of the remaining residual budget, yet raw score differences are small enough that reward shaping and outer-loop hyperparameters must be tuned with the architecture (Section~\ref{sec:reward-shaping}).
Hard-cliff strata make the operative regime explicit: DualCosine drops to $R\approx 0.657$ on the top-10\% wrap-jump tiles while the favorite stays near $0.9915$.
Favorite wrap-jump barely moves toward zero on that subset.
We claim tiled residual repair toward $r^{\star}$, not endpoint closure.

\paragraph{N2N stress test and transfer conditions.}
On the sine{+}cliff holdout, corrupt$\to$corrupt N2N matches the meta favorite within $\approx 10^{-3}$.
Sibling-supervised N2N and the LSTM ceiling sit slightly above.
That is expected for oracle-label training on the same generator and does not invalidate the hybrid search story:
DenoiseOpt discovers compact bake graphs without engine$\to$ideal supervision at search time, and the hard-cliff DualCosine collapse contrast remains.
Win conditions (\texttt{transfer\_failures.json}): favorite beats DualCosine on sine cliffs and most AKWF / Rust tiles; no-bake/FIR lead on exported factory geometry (domain shift, not buried).
On true ReelSynth exports the searched favorite trails no-bake and FIR; we report that gap rather than claim universal factory superiority.

\paragraph{Wrap-jump vs edge RMSE.}
Primary claim remains prolonged $R$. Wrap-jump is diagnostic only.
Edge RMSE is the locked seam-local secondary: favorite reduces it on hard cliffs even when wrap-jump stays flat.
Endpoint-pin zeros wrap-jump while hurting $R$, confirming the objectives differ.

\paragraph{Implication for this paper.}
Mean $R$ (and mean DualCosine gaps) therefore compress a structured difficulty map.
Reporting only the search-campaign mean on synthetic sine cliffs understates how much of the remaining $\sim$1\% sits in a few hard families, and how DualCosine fails when wrap-jump is large.
Hard-cliff strata and wavetable-native transfer (true exports + AKWF) are first-class empirical claims of the present work.
The contribution is aimed at DSP / synthesis artifact repair (DAFx / AES / arXiv cs.SD), not general speech-enhancement leaderboards.

\paragraph{Hybrid search vs.\ plateau.}
Search traces at the $5$k gate show early high champions then long plateaus near $R\approx 0.99$.
That plateau is expected under near-ceiling $R$: further gains are small in absolute units and sensitive to reward shaping, entropy/clip, mutation rates, and plateau-triggered branch reweighting.
Plateau adaptation (deeper cells, denser MoE graphs, higher GA/PBT weight) is the operational response inside one campaign. It does not by itself rebalance family-conditional risk.
A natural next paper is family-conditional and cliff-conditional meta-learning: train or select denoisers per family, or condition the outer loop on measured wrap statistics, and publish per-family residual curves beside the scalar champion---together with ablations of reward shaping under the same compressed-$R$ regime.

\paragraph{Hybrid branches.}
At the reported $5$k-gate search freeze, GA and PBT lead branch-best residuals, with PPO close behind and NAS trailing on this seed (Table~\ref{tab:ablation}).
These per-branch readings are snapshots from a single hybrid run, not causally isolated single-branch experiments.
Isolated short-budget re-runs (150 iterations, same search seed) confirm relative ordering under a fixed compute cap (Table~\ref{tab:ablation}).
The matched $5$k outer-loop comparison (Table~\ref{tab:meta-approaches}, Figures~\ref{fig:meta-approach-bars}--\ref{fig:meta-approach-compare}) places Random NAS, Cont.\ CMA-ES, Arch REINFORCE, Aging evolution, and TPE Bayes NAS beside \textbf{Ours (hybrid GA--PPO)} under a shared seed and bake vocabulary that includes searchable LSTM/xLSTM blocks.
Ours wins the absolute-$R$ gate ($R\approx 0.99146$) with Aging evolution close behind ($R\approx 0.99079$); Cont.\ CMA-ES is the wall-time leader ($\approx 2.3$\,h) at lower champion $R$.
Figure~\ref{fig:meta-heal-samples} shows the wrap-seam heal qualitatively on holdout geometry.
The multi-family SOTA matrix (Table~\ref{tab:sota-main}) shows the meta favorite ahead of DualCosine on $R$, SNR, and SDR across twenty Python generative draws, and ahead of no-bake / FIR on absolute $R$ on that same generator.
Fixed MLP/CNN-on-$R$ baselines beat DualCosine and trail the searched favorite.
On Rust \texttt{sound\_bench} tiles (Table~\ref{tab:rust-bench}) the same favorite beats DualCosine (mean $\Delta R={+}0.063$) and trails no-bake / FIR. That transfer gap matches the narrow Python training geometry.
