\section{Methods}
\label{sec:methods}

\subsection{Notation}
\label{sec:notation}

Throughout, $x\in\mathbb{R}^{L}$ is the cracked single-period wavetable (engine input with open-wrap cliff).
Let $\mathcal{H}$ be the space of bake-cell configurations (discrete operator graph plus continuous bake parameters).
The bake operator is
\begin{equation}
\label{eq:Theta}
\Theta:\mathbb{R}^{L}\times\mathcal{H}\to\mathbb{R}^{L},\qquad
y=\Theta(x;h),\quad h\in\mathcal{H}.
\end{equation}
When continuous parameters alone are emphasized we write $\theta$ for the continuous slice of $h$ and $y=\Theta(x;\theta)$.
$r^{\star}\in\mathbb{R}^{L}$ is the procedural ideal sibling defined below.
Tiling length $N$ (typically $N{=}16$) forms $y_{\mathrm{tiled}}=\mathrm{Tile}(y,N)$ and $r^{\star}_{\mathrm{tiled}}=\mathrm{Tile}(r^{\star},N)$.
Seam width $W$ (code: \texttt{SEAM\_W}, typically $8$) is the neighborhood edited by classical fade and polish ops.
$R\in[0,1]$ is the prolonged residual score ($1$~=~best).
These symbols follow a different convention from Noise2Noise clean/noisy pairs.

\begin{table}[t]
  \centering
  \caption{Core symbols used in Methods.}
  \label{tab:symbols}
  \setlength{\tabcolsep}{3pt}
  \begin{tabular}{@{}cl@{}}
    \toprule
    Symbol & Meaning \\
    \midrule
    $L$ & Cycle length (samples per period) \\
    $x$ & Cracked engine wavetable period (cliff on) \\
    $y=\Theta(x;h)$ & Baked cycle after seam ops / residual cell \\
    $\Theta$ & Bake operator $\mathbb{R}^{L}\times\mathcal{H}\to\mathbb{R}^{L}$ \\
    $h\in\mathcal{H}$ & Bake-cell configuration (ops + continuous $\theta$) \\
    $r^{\star}$ & Ideal sibling (same seed; cliff withheld) \\
    $G$ & Procedural period generator (cliff on/off) \\
    $N$ & Prolong tiling count \\
    $W$ & Seam neighborhood width \\
    $R$ & Prolonged residual score in $[0,1]$ (1~=~best) \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Problem and ideal sibling}
\label{sec:ideal-tile}

\textbf{Problem.} Given a cracked period $x$, choose a bake configuration $h$ so that the repaired cycle $y=\Theta(x;h)$ matches a no-cliff ideal under prolonged tiling.

\textbf{Ideal sibling (Q1).}
Let $G(\mathrm{seed},\mathrm{cliff})$ denote the procedural generator used by the CUDA FitCell batch maker
(\texttt{make\_batch} in \texttt{overnight\_gpu\_rl\_arch.py}): one draw of frequency, phase, and mid-cycle content, with an optional open-wrap cliff of width $W$ at both ends.
For any seed,
\begin{equation}
\label{eq:G-siblings}
r^{\star}=G(\mathrm{seed},\,\mathrm{cliff{=}off}),\qquad
x=G(\mathrm{seed},\,\mathrm{cliff{=}on}).
\end{equation}
Engine and ideal therefore share the mid-cycle content of one draw and differ by the seam cliff (plus any seam-boosted noise applied only on the engine path).
The ideal is \emph{not} a latent studio ground truth and is \emph{not} a method output.
For an arbitrary cracked period from this family, the sibling is obtained by replaying the same seed with the cliff withheld---exactly the pair returned by \texttt{make\_batch}.
This sibling relationship makes unsupervised ranking possible without paired studio recordings~\cite{lehtinen2018,kashyap2021n2n}.

\subsection{Outer objective and residual score $R$}
\label{sec:residual-score}

\textbf{Single outer objective.}
Every candidate bake---including classical controls and searched cells---is ranked by maximizing absolute prolonged residual against the same ideal sibling:
\begin{equation}
\label{eq:outer-obj}
\max_{h\in\mathcal{H}}\; R\bigl(y,\,r^{\star}\bigr)
\quad\text{subject to}\quad
y=\Theta(x;h).
\end{equation}
Equivalently, FitCell minimizes loss $1-R$. Selection, population ranking, and champion updates use absolute $R$ only.

\textbf{Every} scored row---no-bake, DualCosine, FIR, poly, VA residuals, neural favorite---applies some $\Theta$ to the cracked engine $x$, then compares tiled $y$ to the \emph{same} $r^{\star}$ under~\eqref{eq:R}.
The ideal sibling is the scoring target; it is not a method and not the no-bake baseline.

With tiling length $N$,
\begin{equation}
\label{eq:R}
R=\mathrm{clamp}\!\left(1-\frac{\mathrm{rms}(y_{\mathrm{tiled}}-r^{\star}_{\mathrm{tiled}})}{\max(\mathrm{rms}(r^{\star}_{\mathrm{tiled}}),\varepsilon)},\,0,\,1\right),
\end{equation}
where $\mathrm{rms}(z)=\sqrt{\frac{1}{|z|}\sum_i z_i^2}$ and $\varepsilon>0$ avoids division by zero.
By construction $R\in[0,1]$, $R{=}1$ when $y_{\mathrm{tiled}}=r^{\star}_{\mathrm{tiled}}$, and (when pre-clamp scores lie in $(0,1)$) $R$ is strictly decreasing in residual RMS for fixed ideal RMS.
A method ``beats'' another iff its $R$ is higher on the same $(x,r^{\star})$ draw.
Optional soft shape gates (when enabled) multiply $R$ by factors that down-weight mid-cycle damage relative to seam-local error.
We do not claim convergence guarantees for the hybrid outer loop, nor a general wrap-closure theorem for arbitrary bake cells.
Outer-loop cost scales as population size $\times$ proposals per iteration $\times$ fit steps $\times$ batch forward cost (order $10^{3}$--$10^{4}$ arch evaluations at the reported $5$k gate).

\begin{remark}[Noise2Noise motivation]
\label{rem:n2n}
Ranking without studio-clean pairs is possible when the ideal is a procedural sibling of the engine draw~\cite{lehtinen2018,kashyap2021n2n}.
\end{remark}

\subsection{Bake operator $\Theta$ and classical baselines}
\label{sec:bake}

A \textbf{bake cell} $\Theta(\cdot;h)$ composes classical seam ops (DualCosine, polish, soft seam, FIR blends) with optional tiny residual cells (MLP/FIR/U-Net-lite/attention-lite/LSTM-lite) and optional MoE soft gates over experts (Appendix~\ref{app:algorithms}, Algorithms~\ref{alg:dualcosine}--\ref{alg:moe}).
Bake parameters live in a bounded box; architecture strings are discrete graphs over those ops.
Separate classical controls apply cycle-local BLIT/BLEP, PolyBLEP, and BLAMP residual corrections at the wrap on the cracked engine tile (same interface as DualCosine / FIR; not a full oscillator rewrite).
Plain-language roles: BLIT/BLEP and PolyBLEP correct a \emph{value} step at wrap; BLAMP corrects a \emph{slope}/corner jump; DualCosine fades endpoints without a BLEP residual.
They are scored in Section~\ref{sec:va-seam-baselines}.

\paragraph{Four roles (do not conflate).}
\label{sec:four-roles}
\begin{itemize}
  \item \textbf{Ideal sibling} $r^{\star}$: cliff withheld; scoring target only (panel~(a) in Figure~\ref{fig:intro-sine}).
  \item \textbf{No-bake} (passthrough): $\Theta_{\mathrm{no\text{-}bake}}(x)=x$; unrepaired cracked engine scored vs $r^{\star}$ (panel~(b)). Not $r^{\star}$, not a residual-net skip, not a learned identity map.
  \item \textbf{DualCosine}: classical raised-cosine end fade on $x$ (panel~(c)); one classical board row and, separately, a PPO advantage baseline (Remark~\ref{rem:dualcosine-centering}).
  \item \textbf{Ours}: searched bake $y=\Theta(x;h^{\star})$ from the hybrid outer loop (panel~(d)).
\end{itemize}
On mild cliffs residual mass $\ll$ ideal RMS under~\eqref{eq:R}, so no-bake $R$ can sit near $0.97$ while the wrap click remains audible---a ``do nothing'' reference, not a perceptual wrap score.
Released JSON may still use the legacy key \texttt{identity} for no-bake; manuscript prose uses \emph{no-bake}.

\paragraph{Classical (non-AI) comparison board.}
\label{sec:classical-board}
Learned / searched methods are always ranked against the full classical board on absolute $R$ (same $r^{\star}$):
no-bake, DualCosine / cosine fade, \texttt{seam\_fir3}, classic quadratic, hann/crossfade/soft fades, detrend/ensemble, poly seam, and cycle-local BLIT/BLEP / PolyBLEP / BLAMP where scored.
When tables show $\Delta R$ vs DualCosine, that is a reporting gap vs one classical bake, not the definition of reward.

\begin{remark}[DualCosine centering is PPO advantage only (Q2)]
\label{rem:dualcosine-centering}
The outer objective remains~\eqref{eq:outer-obj}: maximize absolute $R(y,r^{\star})$.
In the PPO branch only, we optionally form a centered advantage
\begin{equation}
\label{eq:dualcosine-centering}
\rho = R - R_{\mathrm{DualCosine}},
\end{equation}
so early-search advantages are roughly zero-mean.
This does \emph{not} change selection, FitCell loss, champion ranking, or the matched $5$k meta-approach objective---all of which still maximize absolute $R$.
DualCosine centering is soft relative shaping for the actor--critic, not a hard absolute ranking target.
\end{remark}

\subsection{Pipeline overview}
\label{sec:pipeline-overview}

Figure~\ref{fig:denoiseopt-arch} summarizes the residual-scored hybrid loop before component detail.
Each outer iteration proposes a bake configuration $h$, runs FitCell (Appendix~\ref{app:algorithms}, Algorithm~\ref{alg:fit}) to fit continuous parameters under loss $1-R$, evaluates prolonged $R$ vs $r^{\star}$, and updates the population / PPO policy.
Pseudocode for GA, PPO, PBT, and the full hybrid rotation is in Algorithms~\ref{alg:ga}--\ref{alg:hybrid}.

\input{subsections/arch_diagram}

\subsection{Search space and FitCell}
\label{sec:search-space}

Each genome encodes: ordered op list (fade, polish, FIR, pin, residual write), depth, residual cell type (none, MLP, FIR, U-Net-lite, attn-lite, LSTM-lite), optional MoE expert count $K\in\{2,3,4\}$, and continuous $\theta$ in the CUDA runner box bounds.
The searchable bake vocabulary includes a width-capped 1--2 layer LSTM block over the seam window or full cycle (distinct from the fixed supervised seq-LSTM SOTA baseline).
The live hybrid tag is \texttt{PPO+GA+PBT+NAS+depth+MoE} with LSTM enabled in the block graph.

FitCell draws sibling batches via $G$, applies $\Theta(\cdot;h)$, and steps Adam on $1-R$ with early stopping on plateau (Algorithm~\ref{alg:fit}).

\subsection{Hybrid outer loop}
\label{sec:hybrid}

Each iteration rotates among branches: PPO, GA, PBT, NAS, and combo.
Selection and FitCell are always by absolute residual $R$ (maximize toward the ideal sibling; loss $1-R$).

\paragraph{Near-ceiling $R$ and why reward / search HPs matter.}
\label{sec:reward-shaping}
On the sine{+}cliff generator, no-bake already sits near $R\approx 0.97$, and strong classical / searched bakes live in a narrow band toward $1$ (often $0.99$--$0.991$).
Absolute gaps of order $10^{-3}$--$10^{-2}$ are scientifically real but tiny as raw PPO/GA signals: without shaping, advantages collapse and hyperparameters (clip, entropy, mutation rate, population size, fit steps, learning rate) dominate whether the outer loop can climb the last percent.
We therefore treat \textbf{reward shaping and search hyperparameters as first-class tunables} alongside bake architecture.
In the present hybrid, PPO uses a searchable shaping mode among $\{\texttt{abs\_r},\,\texttt{vs\_dualcosine},\,\texttt{vs\_nobake},\,\texttt{neglog\_gap}\}$ (PBT-mutated; default $\rho = R - R_{\mathrm{DualCosine}}$ per Remark~\ref{rem:dualcosine-centering}).
PBT mutates fit / PPO hyperparameters and reward mode in-population; plateau adaptation retunes branch mix and depth/MoE bias when champions stall.
The matched $5$k meta-approach comparison (Section~\ref{sec:meta-compare}) is the sole GPU experimentation vehicle for outer-loop family, HP, and reward-shaping sweeps under a shared FitCell budget.

\subsection{Meta-learning approach comparison}
\label{sec:meta-compare}

Beyond the hybrid outer loop, we compare five additional outer-loop families under a matched $5{,}000$-evaluation budget, shared search seed \texttt{1902771841}, and the same FitCell / prolonged-$R$ protocol (LSTM and xLSTM included in the searchable bake vocabulary).
Table and figure legends use the short display names below (artifact folders keep stable code keys; see \texttt{NOMENCLATURE.md}):
\begin{itemize}
  \item \textbf{Random NAS} (\texttt{random}): independent uniform draws over bake graphs and fit hyperparameters (control).
  \item \textbf{Cont.\ CMA-ES} (\texttt{cmaes}): diagonal covariance-matrix adaptation over a continuous encoding of depth, width, cell kind, block logits, and fit hyperparameters~\cite{hansen2016cmaes}.
  \item \textbf{Arch REINFORCE} (\texttt{reinforce}): vanilla policy gradient over discrete architecture-edit actions (no clipped surrogate; contrast to PPO~\cite{schulman2017,zoph2017nas}).
  \item \textbf{Aging evolution} (\texttt{aging\_evo}): regularized evolution with oldest-member replacement after tournament parent selection~\cite{real2017large,real2019}, distinct from the hybrid's non-aging tournament GA branch.
  \item \textbf{TPE Bayes NAS} (\texttt{tpe}): lightweight tree-structured Parzen estimator over discrete bake categoricals in the spirit of practical Bayesian optimization~\cite{snoek2012}.
\end{itemize}
We also run \textbf{Ours (hybrid GA--PPO)} (\texttt{hybrid\_lstm}): the multi-branch hybrid loop for $5{,}000$ evaluations with LSTM/xLSTM in the block vocabulary and online reward-mode / fit--PPO hyperparameter co-tuning, as the proposed matched-budget arm.
Results appear in Table~\ref{tab:meta-approaches} and Figures~\ref{fig:meta-approach-bars}--\ref{fig:meta-approach-compare}; healed-sample overlays in Figure~\ref{fig:meta-heal-samples}.

\begin{remark}[Trial complexity]
\label{rem:complexity}
Per outer iteration the dominant cost is $O(T_{\mathrm{fit}}\cdot C_{\mathrm{fwd}})$ for fit steps times one forward bake.
Bake-width caps (tiny residual cells) keep $C_{\mathrm{fwd}}$ far below full-band Demucs/DiffWave forwards screened in Related Work.
\end{remark}

\subsection{Hyperparameters}
\label{sec:hyperparams}

Table~\ref{tab:hyperparams} lists the CUDA search defaults used for the $5$k-gate freeze reported in Results.
Under the near-ceiling residual regime (Section~\ref{sec:reward-shaping}), these outer-loop and fit hyperparameters are co-tuned with architecture; PBT and plateau adaptation continue to mutate several of them online.
Companion listings: Algorithms~\ref{alg:residual}--\ref{alg:hybrid} (Appendix~\ref{app:algorithms}) and \texttt{docs/PSEUDOCODE.md}.

\begin{table}[t]
  \centering
  \small
  \caption{Hybrid search hyperparameters (RTX~3090 DenoiseOpt CUDA search runner / \texttt{denoise\_meta\_evo.py} defaults). Co-tuned with reward shaping under near-ceiling $R$ (Section~\ref{sec:reward-shaping}).}
  \label{tab:hyperparams}
  \setlength{\tabcolsep}{3pt}
  \begin{tabular}{@{}lp{0.48\columnwidth}@{}}
    \toprule
    Parameter & Value \\
    \midrule
    Population size $n$ & 12 \\
    GA tournament $k$ & 3 \\
    GA crossover fraction $p_c$ & $0.5$ (else clone elite parent) \\
    GA mutation & $\ge 1$ mutate after inherit ($+50\%$ second mutate) \\
    PPO clip $\epsilon$ & $0.2$ (default; searchable $\{0.1,0.2,0.25,0.3\}$) \\
    Adam learning rate (fit) & $3{\times}10^{-3}$ (default; PBT-searchable) \\
    Adam learning rate (policy) & $3{\times}10^{-4}$ \\
    Entropy coefficient (PPO) & $0.02$ (default; searchable $\approx 0.005$--$0.06$) \\
    PBT exploit / mutate & elite fraction $0.25$; multi-step arch+hp mutate \\
    MoE routing & \texttt{sequential} or \texttt{moe\_parallel} when enabled \\
    Plateau boredom threshold & $1000$ iters without champ improve \\
    Cycle length $L$ / tiles $N$ / seam $W$ & $256$ / $16$ / $8$ \\
    Fit steps / batch (default) & $24$ / $48$ \\
    Algo tag & \texttt{PPO+GA+PBT+NAS+depth+MoE} \\
    \bottomrule
  \end{tabular}
\end{table}
