\section{Tooling and AI use}
\label{sec:tooling}

This section discloses research and writing tools, including AI systems, with
scope and intent.

\paragraph{Literature discovery.}
Open literature was gathered with the Klaut Research Gateway (local OpenAlex /
Crossref / arXiv providers) under queries spanning virtual-analog anti-aliasing,
unsupervised restoration, HPO/PBT, RL--GA hybrids for NAS, and audio denoise
architectures.
Harvest JSONs live in the companion meta repository.
Source PDFs were downloaded for every bibliography entry
(\texttt{artifacts/literature\_oa/pdfs/}).
Paywalled-only items were dropped or replaced.

\paragraph{Paper drafting.}
Section planning, subsection drafts, and revision passes used Klaut
\texttt{research\_paper\_*} tools with local Ollama (\texttt{qwen3.5:9b}) when
available, under an explicit scientific workflow (plan $\rightarrow$ write
$\rightarrow$ revise $\rightarrow$ export) into an arXiv-style twocolumn
template.
AI drafts were treated as scaffolding. Claims, numbers, and citations
were checked against harvest records, PDF analyses, and measured artifacts.
Final prose was edited for DSP terminology and to remove generic filler.

\paragraph{Why AI was used.}
(1)~Scale literature triage across DSP and AutoML without pretending every
high-cite hit is in-domain.
(2)~Accelerate IMRaD scaffolding so human effort
focuses on metric honesty (residual vs DualCosine, family hardness, gate vs
final $5$k-gate search claims).
(3)~Keep an auditable tool chain (plans, PDF
snippet analyses, citation flags) over opaque chat-only writing.

\paragraph{Where AI was \emph{not} used.}
Model fitting, residual scoring, DualCosine baselines, lit-combo / hybrid
GPU search, and inference-time benchmarks are conventional code (Rust / PyTorch)
with deterministic logs and JSON artifacts.
AI did not invent experimental numbers.

\paragraph{Other tools.}
Rust/\texttt{cargo} for the procedural bench and meta binaries. PyTorch CUDA
for hybrid GA--PPO search. matplotlib for figures. MiKTeX \texttt{pdflatex}
for PDF builds. GitHub for versioned papers and artifacts.

\paragraph{Hear samples and vibrato figures.}
Audible wrap demos: \texttt{reelsynth/brand/artifacts/meta\_approach\_compare/hear\_samples/}
(rebuild: \texttt{scripts/export\_meta\_hear\_samples.py}).
Vibrato spectrogram bench: \texttt{scripts/bench\_vibrato\_spectrogram.py}
$\to$ \texttt{brand/artifacts/vibrato\_spectrogram/} (mirrored into \texttt{paper/v8/figures/}).
Listening panels for the manuscript are spectrogram + downloadable WAV citations only.
